Hyperplane Elimination for Quickly Enumerating Local Optima

### 2015-09-22 Hyperplane Enumeration/main.tex
Updated abstract from writing group

\begin{abstract}
A common tool for understanding combinatorial problem landscapes
is to examine the properties of its local optima. Unfortunately,
exhaustive methods for finding local optima are exceptionally time
consuming and are therefore limited to very small problem sizes.
We propose a method for exploiting generalized problem structure
to significantly speed up this process by skipping hyperplanes
which cannot contain local optima. In most of our experiments
this hyperplane elimination allows enumeration time to scale with
the total number of local optima instead of the total landscape size.
We prove optimality for linear and $k$-bounded problems, and provide
empirical evidence of optimality on NKq Landscapes and Ising Spin Glasses.
Furthermore, we are able to extend this method to finding only solutions
which cannot be improved by flipping $r$ or fewer bits. In some cases
increasing $r$ actually reduces total runtime by drastically reducing the
number of local optima. While previous methods were limited to
problems with at most 18 or 30 bits, hyperplane elimination can enumerate the same problems with
77 bits. When using higher $r$ values this bound can be pushed to 200 bit problems
on average enumerated in under 6 minutes.

>>>>

\begin{abstract}
Examining the properties of local optima is a common method for understanding
combinatorial-problem landscapes.
Unfortunately,
exhaustive methods for finding local optima are limited to very small problem sizes.
We propose a method for exploiting generalized problem structure
to skip hyperplanes
that cannot contain local optima. In most of our experiments,
this hyperplane elimination allows enumeration time to scale with
the total number of local optima instead of the total landscape size.
We prove optimality for linear and $k$-bounded problems, and we provide
empirical evidence of optimality on NKq Landscapes and Ising Spin Glasses.
Furthermore, we are able to refine this method to find only solutions
that cannot be improved by flipping $r$ or fewer bits. Counterintuitively,
increasing $r$ can reduce total runtime by drastically reducing the
number of local optima. While previous methods were limited to
landscapes with at most $2^{30}$ binary strings, hyperplane elimination can enumerate the same problems with
$2^{77}$ binary strings, and find all 4-bit local optima to problems with $2^{200}$ binary strings.
We believe this increase in size will help further our understanding of complex real-world landscapes.
